### 2021-09-21 papers/modeling-of-combined-lfr-csp-sco2-paper/content.tex
BL commits on results & discussion

Mixing two different temperature flows before the turbine in a Brayton cycle has a negative effect on cycle efficiency \mw{compared to...?}. To quantify the reduction in cycle efficiency, another cycle with no mixing prior to the turbine is desired. This cycle, C-2HTR3T-ON, can be seen in Figure \ref{c-2htr3t-on} and has two high temperature recuperators and three turbines. The LFR is powering one turbine, T1, and transferring unused heat to the flow entering LFR HX through a dedicated high temperature recuperator, \hl{HTR1}. 
%\mw{The LFR doesn't recuperate heat. Clarify what you mean here. -BW} 
The cycle with heat addition from the CSP also contains two separate turbines, T2, while having a dedicated high temperature recuperator, HTR2. 
\mw{Need to explain how T2 is actually 2 separate turbines represented diagramatically as 1, and why we can model it this way.} 
After the high temperature recuperators, the two flows are combined and sent to the LTR hot side. 
%\bl{this is more about having a compromise option of intermediate complexity and efficiency}

@@

The two-cycle configuration has the CSP and LFR operating in separate recompression cycles when the focus is on electrical generation, therefore these two cycles are analyzed individually. C-LFR-ON has the highest efficiency during unconstrained operation with a gain of 1.8\%. The unconstrained case allows for smaller temperature gradient across the LFR, an increase in recuperator effectiveness, and an increase in recuperator capacitance ratio, all of which reduce sources of irreversibility in the cycle \cite{klein_nellis_2011}. According to studies carried out by our industry partner, Westinghouse, increasing the LFR sCO$_2$ inlet temperature above the design value of 400$^{\circ}$C has a negative effect on LFR HX outlet temperature, effectively demanding an ideal LFR heat exchanger, but does increase the LFR efficiency. Additionally, Westinghouse concluded that a lower sCO$_2$ LFR inlet temperature, down to 340$^{\circ}$C, is desirable for higher core power. The sCO$_2$ inlet temperature to the LFR increases by 15.1$^{\circ}$C to a value of 415.1$^{\circ}$C, increasing LFR efficiency but demanding a larger, more idealized, LFR heat exchanger.   
\hl{would I need more information on the LFR from westinghouse (ie report) to make these statements? additionally a conclusion to these statements needs more supporting detail, should I go with the higher efficiency or the cycle that I know works with design values?}

C-CSP-ON is not affected by constrained or unconstrained LFR sCO$_2$ inlet temperature because the two cycles are not tethered during non-charging configuration. Two individual studies are conducted on the C-CSP-ON cycle cold TES temperature, one with a lower design value of 390$^{\circ}$C and another with a higher value of 440$^{\circ}$C. A cold TES temperature of 410$^{\circ}$C is excluded since it can be presumed to have an efficiency intermediate to the 390$^{\circ}$C and 440$^{\circ}$C calculated values. Cold TES temperature is found to have a negligible effect on cycle efficiency between the two studied cases, with the 440$^{\circ}$C case gaining 0.35\%. The component that has the largest change in values is the CSP HX with a large drop in capacitance ratio and UA value, therefore the lower temperature of 390$^{\circ}$C for cold TES temperature is ideal increasing the performances of the heat exchanger.

\subsubsection{C-1HTR1T-ON}
The single cycle configuration, C-1HTR1T-ON, has the CSP and LFR heat additions in parallel causing identical inlet conditions. Due to the 10$^{\circ}$C approach temperature, the identical inlet conditions lead to a lower bound of 410$^{\circ}$C on the cold TES temperature when the sCO$_2$ LFR inlet temperature is constrained to 400$^{\circ}$C. A supplementary 440$^{\circ}$C cold TES study is ran to test the effects that higher temperature storage has on cycle efficiency. Raising the temperature of the cold TES from 410$^{\circ}$C to 440$^{\circ}$C has no effect on cycle efficiency but did raise the solar salt mass flow rate to accommodate for the lower temperature difference across the CSP HX hot side. Increasing the solar salt mass flow rate requires more storage for the same CSP heat input, a reduction in dispatchability, and larger pumps, increasing the cost of the system. Therefore, a lower cold TES temperature is desirable. Without the constraint on the sCO$_2$ LFR inlet temperature, a study is additionally conducted with cold TES temperature set to 390$^{\circ}$C. Of these three sensitivity studies, the cycle that yielded the highest efficiency is the unconstrained, 390$^{\circ}$C cold TES with an efficiency gain of 0.68\% over the constrained cases. The high temperature difference between the hot and cold TES reduces the mass flow rate of solar salt and subsequent cost of system components. The sCO$_2$ LFR inlet temperature for the 390$^{\circ}$C cold TES study is 380$^{\circ}$C, which increases the performance of the LFR heat exchanger as well as allowing for a higher core power.

@@

C-LFR-CIRC is a thermal energy storage charging technique which does not suffer from losses associated with turbines, compressors, and heat exchangers. This is due to the fact that a separate recirculation cycle is modeled taking heat directly from the LFR, passing it through a recirculating pump, exchanges the heat into the hot TES through C2S, then removes heat in a chiller to bring the temperature of the flow down to the inlet temperature of the LFR, 400$^{\circ}$C. This is the most adoptable TES charging technique a wide range of possible TES charging temperatures with almost all of the heat from the LFR being transferred into the hot TES depending on the temperature. Multiple temperatures for the cold TES are studied; 390$^{\circ}$C, 410$^{\circ}$C, and 440$^{\circ}$C. Of these three temperatures, the highest efficiency is a cold TES temperature of 390$^{\circ}$ with a heat storage efficiency of 99.92\%. The heat storage efficiency reduces as the temperature of the cold TES is increased, the 410$^{\circ}$C case has an efficiency of 89.66\% while the 440$^{\circ}$C case has an efficiency of 74.29\%. This reduction in efficiency is due to the inlet temperature condition on the LFR, the higher the temperature of the cold TES the higher the temperature of the C2S outlet and therefore the more heat is being extracted through the chiller to bring the flow temperature down to the LFR inlet temperature. A drawback of this TES charging technique is the additional components needed for the circulation loop, with an additional chiller and circulator component, adding cost.

>>>>

Mixing two different temperature flows before the turbine in a Brayton cycle has a negative effect on cycle efficiency \mw{compared to...?}. To quantify the reduction in cycle efficiency, another cycle with no mixing prior to the turbine is desired. This cycle, C-2HTR3T-ON, can be seen in Figure \ref{c-2htr3t-on} and has two high temperature recuperators and three turbines. The LFR is powering one turbine, T1, and transferring unused heat to the flow entering LFR HX through a dedicated high temperature recuperator, \hl{HTR1}. 
%\mw{The LFR doesn't recuperate heat. Clarify what you mean here. -BW} 
The cycle with heat addition from the CSP also contains two separate turbines, T2, while having a dedicated high temperature recuperator, HTR2. 
\mw{Need to explain how T2 is actually 2 separate turbines represented diagramatically as 1, and why we can model it this way.} 
After the high temperature recuperators, the two flows are combined and sent to the LTR hot side. 
%\bl{this is more about having a compromise option of intermediate complexity and efficiency}
\bl{also why this isnt really an addition turbine, it is using the turbines that power the compressors}

@@

The two-cycle configuration has the CSP and LFR operating in separate recompression cycles when the focus is on electrical generation, therefore these two cycles are analyzed individually. C-LFR-ON has the highest efficiency during unconstrained operation with a gain of 1.8\%. The unconstrained case allows for smaller temperature gradient across the LFR, an increase in recuperator effectiveness, and an increase in recuperator capacitance ratio, all of which reduce sources of irreversibility in the cycle \cite{klein_nellis_2011}. Increasing the LFR sCO$_2$ inlet temperature above the design value of 400$^{\circ}$C is favorable from a cycle efficency perspective, but reduces the LFR thermal power. This is because the LFR mass flow rate and outlet temperature are constrained by materials properties (to limit creep, corrosion and erosion), so a lower inlet temperature implies a lower thermal power. In principle, dropping the LFR inlet temperature further, to as low as 340^{\circ}$C (beyond which lead freezing becomes limiting), would further increase thermal power at the expense of cycle efficiency, so a trade-off is needed. When unconstrained, the sCO$_2$ inlet temperature to the LFR increases by 15.1$^{\circ}$C to a value of 415.1$^{\circ}$C, increasing LFR efficiency but demanding a larger, more idealized, LFR heat exchanger.   
\bl{text revised}

C-CSP-ON is not affected by constrained or unconstrained LFR sCO$_2$ inlet temperature because the two cycles are not tethered during non-charging configuration. \bl{Provided that the charging configuration is suitably configured} Two individual studies are conducted on the C-CSP-ON cycle cold TES temperature, one with a lower design value of 390$^{\circ}$C and another with a higher value of 440$^{\circ}$C. A cold TES temperature of 410$^{\circ}$C is excluded since it can be presumed to have an efficiency intermediate to the 390$^{\circ}$C and 440$^{\circ}$C calculated values. Cold TES temperature is found to have a negligible effect on cycle efficiency between the two studied cases, with the 440$^{\circ}$C case gaining 0.35\%. The component that has the largest change in values is the CSP HX with a large drop in capacitance ratio and UA value, therefore the lower temperature of 390$^{\circ}$C for cold TES temperature is ideal increasing the performances of the heat exchanger. 
\bl{can you produce a combined efficiency which is total cycle electrical output/total heat input for C-LFR-ON + C-CSP-ON in combination}

\subsubsection{C-1HTR1T-ON}
The single cycle configuration, C-1HTR1T-ON, has the CSP and LFR heat additions in parallel causing identical inlet conditions. Due to the 10$^{\circ}$C approach temperature, the identical inlet conditions lead to a lower bound of 410$^{\circ}$C on the cold TES temperature when the sCO$_2$ LFR inlet temperature is constrained to 400$^{\circ}$C. A supplementary 440$^{\circ}$C cold TES study is run to test the effects that higher temperature storage has on cycle efficiency. Raising the temperature of the cold TES from 410$^{\circ}$C to 440$^{\circ}$C has no effect on cycle efficiency but did raise the solar salt mass flow rate to accommodate for the lower temperature difference across the CSP HX hot side. Increasing the solar salt mass flow rate requires more storage for the same CSP heat input, a reduction in dispatchability, and larger pumps, increasing the cost of the system. Therefore, a lower cold TES temperature is desirable. Without the constraint on the sCO$_2$ LFR inlet temperature, a study is additionally conducted with cold TES temperature set to 390$^{\circ}$C. Of these three sensitivity studies, the cycle that yielded the highest efficiency is the unconstrained, 390$^{\circ}$C cold TES with an efficiency gain of 0.68\% over the constrained cases. The high temperature difference between the hot and cold TES reduces the mass flow rate of solar salt and subsequent cost of system components. The sCO$_2$ LFR inlet temperature for the 390$^{\circ}$C cold TES study is 380$^{\circ}$C, which increases the performance of the LFR heat exchanger as well as allowing for a higher core power.

@@

C-LFR-CIRC is a thermal energy storage charging technique which does not suffer from significant losses associated with turbines, compressors, and heat exchangers. This is due to the fact that a separate recirculation cycle is modeled taking heat directly from the LFR, passing it through a recirculating pump, exchanges the heat into the hot TES through C2S, then removes heat in a chiller to bring the temperature of the flow down to the inlet temperature of the LFR, 400$^{\circ}$C. This is the most adoptable TES charging technique a wide range of possible TES charging temperatures with almost all of the heat from the LFR being transferred into the hot TES depending on the temperature. Multiple temperatures for the cold TES are studied; 390$^{\circ}$C, 410$^{\circ}$C, and 440$^{\circ}$C. Of these three temperatures, the highest efficiency is a cold TES temperature of 390$^{\circ}$ with a heat storage efficiency of 100\%. The heat storage efficiency reduces as the temperature of the cold TES is increased, the 410$^{\circ}$C case has an efficiency of 90\% while the 440$^{\circ}$C case has an efficiency of 74\%. This reduction in efficiency is due to the inlet temperature condition on the LFR, the higher the temperature of the cold TES the higher the temperature of the C2S outlet and therefore the more heat is being extracted through the chiller to bring the flow temperature down to the LFR inlet temperature. A drawback of this TES charging technique is the additional components needed for the circulation loop, with an additional chiller and circulator component, adding cost. 
\bl{need to comment on how this interacts with the choice of cycle configuration. If the LFR temperature is 400C and cold TES is 390C then CSP inlet must be 380C and this requires 2HTR3T or separate cycles. If 1HTR1T is used, then 410C+ is needed which implies an efficiency penalty}
